Soss: Declarative Probabilistic Programming via Runtime Code Generation

### 2020-01-10 paper/paper.tex
Add Taine's email
edits

\author{Taine Zhao}
%\orcid{1234-5678-9012-3456}
\affiliation{%
  \institution{University of Tsukuba}
  \department{Department of Computer Science}
  %\streetaddress{625 Mt Auburn St #3}
  %\city{Cambridge}
  %\state{MA}
  %\postcode{02138}
  %\country{USA}
}
% \email{apfeffer@cra.com}
%\renewcommand\shortauthors{Mage, M. et al}

@@

\begin{description}  
    \item[\texttt{rand}] draws a random point or sample from a given distribution.
    \item[\texttt{particles}] uses \jl#MonteCarloMeasurements.jl# to draw a \emph{systematic sample} \cite{Douc2005}.
    \item[\texttt{logpdf}] takes a distribution and a point, and returns the log-density at that point.
    \item[\texttt{xform}] takes a distribution and some known data, and returns a bijection from $\mathbb{R}^n$ to the space of unknowns. This is useful for algorithms like Hamiltonian Monte Carlo \cite{Neal2011}.
    \item[\texttt{symlogpdf}] takes a model, and returns a SymPy \cite{10.7717/peerj-cs.103} representation of the log-density, with some additional Julia-based simplifications.
    \item[\texttt{codegen}] calls \jl#symlogpdf#, and uses the result to generate a more efficient version of \jl#logpdf#.
\end{description}

@@

\emph{Turing} \cite{ge2018t} is syntactically similar to Soss, but favors a more imperative style. Turing can represent a wider variety of models than Soss, in particular Bayesian nonparameteric models, but sacrifices some ability to reason staticically about models in order to achieve this. We hope to soon find ways to interface with Turing so users can benefit from both Turing's modeling flexibility and Soss's static analysis and specialized code generation.

>>>>

\author{Taine Zhao}
%\orcid{1234-5678-9012-3456}
\affiliation{%
  \institution{University of Tsukuba}
  \department{Department of Computer Science}
  %\streetaddress{625 Mt Auburn St #3}
  %\city{Cambridge} 
  %\state{MA}
  %\postcode{02138}
  %\country{USA}
}
\email{thaut@logic.cs.tsukuba.ac.jp}
%\renewcommand\shortauthors{Mage, M. et al}

@@

\begin{description}  
    \item[\texttt{rand}] draws a random point or sample from a given distribution.
    \item[\texttt{particles}] draws a \emph{systematic sample} using \jl#MonteCarloMeasurements.jl# \cite{8ff6a743-0ad6-4d98-bbb3-5d549c698bc1}.
    \item[\texttt{logpdf}] takes a distribution and a point, and returns the log-density at that point.
    \item[\texttt{xform}] takes a distribution and some known data, and returns a bijection from $\mathbb{R}^n$ to the space of unknowns. This is useful for algorithms like Hamiltonian Monte Carlo \cite{Neal2011}.
    \item[\texttt{symlogpdf}] takes a model, and returns a SymPy \cite{10.7717/peerj-cs.103} representation of the log-density, with some additional Julia-based simplifications.
    \item[\texttt{codegen}] calls \jl#symlogpdf#, and uses the result to generate a more efficient version of \jl#logpdf#.
\end{description}

@@

\emph{Turing} \cite{ge2018t} is syntactically similar to Soss, but tends to favor a more imperative style. Turing calls the Julia compiler at a much earlier stage than Soss. This allows representation of a wider variety of models, at the cost of some flexibility of model transformations. We hope to soon find ways to interface with Turing so users can benefit from both Turing's modeling flexibility and Soss's static analysis and specialized code generation.
